Do I Need to Model You in order to Learn With You?

### 2015-04-21 mutual-modeling.tex
Completed citations

From its very beginning, CSCL research has been following
\citet{roschelle1995construction} suggestion that collaborative learning has
something to do with the process of constructing and maintaining a \emph{shared
understanding} of the task at hand. Building a shared/mutual understanding
refers to the upper class of collaborative learning situations, those in which
students should build upon each other's understanding to refine their own
understanding: what is expected to produce learning is not the mere fact that
two students build the same understanding but the cognitive effort they have to
engage to build this shared understanding~\citep{schwartz1995emergence}. This
effort can be observed by the frequency of rich interactions, \ie interactions
whose occurrence has been related to learning: (self-) explanations in cognitive
science \fixme{REF:Chi} \fixme{REF: Webb}, conflict resolution in
socio-cognitive theories \fixme{REF:Doise},
mutual regulation \fixme{REF: Blaye} in a Vygostkian perspective, etc. The
construction of a shared understanding has been investigated for several years
in psycholinguistics, under the  notion of
\emph{grounding}~\citep{clark1986referring}. However, the relevance of grounding
mechanisms for explaining learning outcomes has been questioned in learning
sciences. The monitoring and repair of misunderstanding explains for instance
referential failures in short dialogue episodes but does hardly predict
\emph{conceptual change} (\ie the acquisition, acceptation and integration of a
new belief into one's mental model\fxnote{check with Pierre}) over longer
sessions~\cite{dillenbourg2006sharing}. The cumulative effect of grounding
episodes can probably be better understood from a socio-cultural perspective:

@@

Mutual modelling does not occur in a vacuum but it is highly contextualized.
REF\fxfatal{REF missing} reviews how the features of the collaborative
situation, namely the media (co-temporality,\ldots), may facilitate or hamper
mutual modelling.  \citet{hutchins1997constructing} reported a study in which a
short silence between two pilots was perfectly interpreted because it occurred
in a highly constrained communication context.  Some environments are more
productive than others in helping peers to detect their misunderstandings.
\citet{roschelle1995construction} reformulate the design of CSCL
interfaces to provide ways for peers to detect and repair their
misunderstanding. In CSCW, researchers have explored various so-called
``awareness tools'', \ie functionalities that inform $A$ is about $B$'s actions
that A could not directly perceive because $B$ was working on a different subset
of the virtual space. Different awareness tools are presented and relied upon in
this contribution as well since they allowed us to manipulate experimentally the
mutual modelling activity.

@@

\paragraph{Study-specific questions} Our hypothesis about media richness is
rejected: the average \Model{A}{B}{emotions} was .62 ($SD = .28$) with the
audio+video condition and .71 ($SD = .19$) in the audio condition ($F=3.4, p:
0.065$). Our interpretation is that viewing what one's partner sees (shared
graphical editor) was, in this task, more important than seeing each
other~\citep{gaver1993one}\fixme{check ref with Pierre}\fixme{other cite:
Anderson et al., 2000}.

>>>>

From its very beginning, CSCL research has been following
\citet{roschelle1995construction} suggestion that collaborative learning has
something to do with the process of constructing and maintaining a \emph{shared
understanding} of the task at hand. Building a shared/mutual understanding
refers to the upper class of collaborative learning situations, those in which
students should build upon each other's understanding to refine their own
understanding: what is expected to produce learning is not the mere fact that
two students build the same understanding but the cognitive effort they have to
engage to build this shared understanding~\citep{schwartz1995emergence}. This
effort can be observed by the frequency of rich interactions, \ie interactions
whose occurrence has been related to learning: (self-) explanations in cognitive
science~\cite{chi1989self, webb1991task}, conflict resolution in
socio-cognitive theories~\cite{doise1975social},
mutual regulation~\cite{blaye1995collaborative} in a Vygostkian perspective, etc. The
construction of a shared understanding has been investigated for several years
in psycholinguistics, under the  notion of
\emph{grounding}~\citep{clark1986referring}. However, the relevance of grounding
mechanisms for explaining learning outcomes has been questioned in learning
sciences. The monitoring and repair of misunderstanding explains for instance
referential failures in short dialogue episodes but does hardly predict
\emph{conceptual change} (\ie the acquisition, acceptation and integration of a
new belief into one's mental model\fxnote{check with Pierre}) over longer
sessions~\cite{dillenbourg2006sharing}. The cumulative effect of grounding
episodes can probably be better understood from a socio-cultural perspective:

@@

Mutual modelling does not occur in a vacuum but it is highly contextualized.
\citet{clark1991grounding} review how the features of the collaborative
situation, namely the media (co-temporality,\ldots), may facilitate or hamper
mutual modelling.  \citet{hutchins1997constructing} reported a study in which a
short silence between two pilots was perfectly interpreted because it occurred
in a highly constrained communication context.  Some environments are more
productive than others in helping peers to detect their misunderstandings.
\citet{roschelle1995construction} reformulate the design of CSCL
interfaces to provide ways for peers to detect and repair their
misunderstanding. In CSCW, researchers have explored various so-called
``awareness tools'', \ie functionalities that inform $A$ is about $B$'s actions
that A could not directly perceive because $B$ was working on a different subset
of the virtual space. Different awareness tools are presented and relied upon in
this contribution as well since they allowed us to manipulate experimentally the
mutual modelling activity.

@@

\paragraph{Study-specific questions} Our hypothesis about media richness is
rejected: the average \Model{A}{B}{emotions} was .62 ($SD = .28$) with the
audio+video condition and .71 ($SD = .19$) in the audio condition ($F=3.4, p:
0.065$). Our interpretation is that viewing what one's partner sees (shared
graphical editor) was, in this task, more important than seeing each
other~\citep{gaver1993one,anderson1997impact}.
